---
title: 'Mario Kart Dashboard Project Part 1'
author: Eric Collins
date: '2022-02-07'
slug: []
categories: []
tags:
comments: no
showcomments: yes
showpagemeta: yes
---


I was watching Youtube videos (as I am doing so often), and came across [this](https://www.youtube.com/watch?v=xo-Qt2mkjQs&t=486s) video from a relatively small Mario Kart Youtuber named Bayesic. He describes in great detail how player stats in Mario Kart 8 Deluxe are calculated and used in game, and how to optimize a build and strategy through the lens of Pareto optimality. He builds a dashboard to optimize any given set of stats. I thought that was a great idea, and sounded like a good way to learn web scraping as well as a new (to me) shiny framework called shinydashboards.


To start with, let's pull the tables from the internet. The website can be found [here](https://www.mariowiki.com/Mario_Kart_8_Deluxe_in-game_statistics)

I'll be using the rvest package to translate the html into tables that can be saved into a feather.

```{r}
library(rvest)
library(tidyverse)
library(janitor)
library(feather)
```


```{r}
url <- "https://www.mariowiki.com/Mario_Kart_8_Deluxe_in-game_statistics"

kart_page <- read_html(url)

str(kart_page)
```


The rvest package uses node syntax to describe what's happening inside the html. To get deeper into the html, keep calling html_children

```{r}
nodes <- kart_page %>%
        html_elements("body") %>%
        html_children()

nodes %>%
        html_children()
```


The tables come out not very pretty. With some custom functions and lapply, we can change that.

```{r}
kart_tables <- kart_page %>%
        html_table(fill = TRUE)

kart_tables1 <- kart_tables[c(2:10)]
kart_tables2 <- kart_tables[c(11:12)]

fix_names_1 <- function(x){
        new_x <- x %>%
                row_to_names(row_number = 1) %>%
                clean_names()
        
        dont <- names(new_x)[1]
        
        new_x <- new_x %>%
                mutate(across(!dont, as.numeric))
}

fix_names_2 <- function(x){
        new_x <- x %>%
                row_to_names(row_number = 2) %>%
                clean_names()
        
        dont <- names(new_x)[1]
        
        new_x <- new_x %>%
                mutate(across(!dont, as.numeric))
}

kart_tables_fixed1 <- lapply(kart_tables1, fix_names_1)
kart_tables_fixed2 <- lapply(kart_tables2, fix_names_2)


kart_tables_fixed <- append(kart_tables_fixed1, kart_tables_fixed2)
```

we can't tibble a list, but we can tibble a dataframe from a list. They are then written to csv and feather, for data longetivity and for fastest loading speed. 

```{r}

karts <- as_tibble(as.data.frame(kart_tables_fixed[1]))
characters <- as_tibble(as.data.frame(kart_tables_fixed[2]))
tires <- as_tibble(as.data.frame(kart_tables_fixed[3]))
gliders <- as_tibble(as.data.frame(kart_tables_fixed[4]))
kart_mass <- as_tibble(as.data.frame(kart_tables_fixed[5]))
acceleration <- as_tibble(as.data.frame(kart_tables_fixed[6]))
slip <- as_tibble(as.data.frame(kart_tables_fixed[7]))
offroad_brake <- as_tibble(as.data.frame(kart_tables[9])) %>%
        slice(1:22) %>%
        row_to_names(row_number = 1) %>%
        clean_names()
        
offroad_slip <- as_tibble(as.data.frame(kart_tables[9])) %>%
        slice(23:45) %>%
        row_to_names(row_number = 1) %>%
        clean_names() 

boost <- as_tibble(as.data.frame(kart_tables_fixed[9]))
speed <- as_tibble(as.data.frame(kart_tables_fixed[10]))
handling <- as_tibble(as.data.frame(kart_tables_fixed[11]))


write_csv(karts, "karts.csv")
write_csv(characters, "characters.csv")
write_csv(tires, "tires.csv")
write_csv(gliders, "gliders.csv")
write_csv(kart_mass, "kart_mass.csv")
write_csv(acceleration, "acceleration.csv")
write_csv(slip, "slip.csv")
write_csv(offroad_brake, "offroad_brake.csv")
write_csv(offroad_slip, "offroad_slip.csv")
write_csv(boost, "boost.csv")
write_csv(speed, "speed.csv")
write_csv(handling, "handling.csv")



write_feather(karts, "karts.feather")
write_feather(characters, "characters.feather")
write_feather(tires, "tires.feather")
write_feather(gliders, "gliders.feather")
write_feather(kart_mass, "kart_mass.feather")
write_feather(acceleration, "acceleration.feather")
write_feather(slip, "slip.feather")
write_feather(offroad_brake, "offroad_brake.feather")
write_feather(offroad_slip, "offroad_slip.feather")
write_feather(boost, "boost.feather")
write_feather(speed, "speed.feather")
write_feather(handling, "handling.feather")
```




















I'm posting this from the future.

The app can be found [here](https://eric-collins.shinyapps.io/MarioKartBuilder/?_ga=2.256633611.166118430.1644261128-1683136978.1644261128).

The github repo can be found [here](https://github.com/eric-collins/MK8D_Stat_Optimizer). 














